---
title: "Homework 4: Machine Learning"
author: VAANI KOHLI
date: 01/06/2023
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
  pdf_document:
    toc: yes
---

```{r}
#| label: load-libraries
#| echo: false # This option disables the printing of code (only output is displayed).
#| message: false
#| warning: false
options(scipen = 999) #disable scientific notation
library(tidyverse)
library(tidymodels)
library(GGally)
library(sf)
library(leaflet)
library(janitor)
library(rpart.plot)
library(here)
library(scales)
library(vip)
```

# The Bechdel Test

<https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-hollywoods-exclusion-of-women/>

The [Bechdel test](https://bechdeltest.com) is a way to assess how women are depicted in Hollywood movies. In order for a movie to pass the test:

1.  It has to have at least two [named] women in it
2.  Who talk to each other
3.  About something besides a man

There is a nice article and analysis you can find here <https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-hollywoods-exclusion-of-women/> We have a sample of 1394 movies and we want to fit a model to predict whether a film passes the test or not.

```{r read_data}

bechdel <- read_csv(here::here("data", "bechdel.csv")) %>% 
  mutate(test = factor(test)) 
glimpse(bechdel)

```

How many films fail/pass the test, both as a number and as a %?

```{r}
bechdel %>%
  filter(test == "Pass") %>%
  nrow()

percentage_pass <- (nrow(subset(bechdel, test == "Pass"))/nrow(bechdel)) * 100

bechdel %>%
  filter(test == "Fail") %>%
  nrow()

percentage_fail <- (nrow(subset(bechdel, test == "Fail"))/nrow(bechdel)) * 100

# 622 films passed, this is 44.6% of all films. 772 films failed, this is 55.3% of all films. 
```

## Movie scores

```{r}
ggplot(data = bechdel, aes(
  x = metascore,
  y = imdb_rating,
  colour = test
)) +
  geom_point(alpha = .3, size = 3) +
  scale_colour_manual(values = c("tomato", "olivedrab")) +
  labs(
    x = "Metacritic score",
    y = "IMDB rating",
    colour = "Bechdel test"
  ) +
 theme_light()
```

# Split the data

```{r}
# **Split the data**

set.seed(123)

data_split <- initial_split(bechdel, # updated data
                           prop = 0.8, 
                           strata = test)

bechdel_train <- training(data_split) 
bechdel_test <- testing(data_split)
```

Check the counts and % (proportions) of the `test` variable in each set.

```{r}
bechdel_train %>%
  filter(test == "Pass") %>%
  nrow()

percentage_pass_train <- (nrow(subset(bechdel, test == "Pass"))/nrow(bechdel)) * 100

bechdel_train %>%
  filter(test == "Fail") %>%
  nrow()

percentage_fail_train <- (nrow(subset(bechdel, test == "Fail"))/nrow(bechdel)) * 100

# In the train data set, 497 or 44.6% films passed and 617 or 55.3% films failed.

bechdel_test %>%
  filter(test == "Pass") %>%
  nrow()

percentage_pass_test <- (nrow(subset(bechdel, test == "Pass"))/nrow(bechdel)) * 100

bechdel_test %>%
  filter(test == "Fail") %>%
  nrow()

percentage_fail_test <- (nrow(subset(bechdel, test == "Fail"))/nrow(bechdel)) * 100

# In the test data set, 125 or 44.6% films passed and 155 or 55.3% films failed.
```

## Feature exploration

## Any outliers?

```{r}

bechdel %>% 
  select(test, budget_2013, domgross_2013, intgross_2013, imdb_rating, metascore) %>% 

    pivot_longer(cols = 2:6,
               names_to = "feature",
               values_to = "value") %>% 
  ggplot()+
  aes(x=test, y = value, fill = test)+
  coord_flip()+
  geom_boxplot()+
  facet_wrap(~feature, scales = "free")+
  theme_bw()+
  theme(legend.position = "none")+
  labs(x=NULL,y = NULL)

# There is 1 film that failed the Bechdel test that seems to have a higher budget that the rest. 2 films that passed and 1 that failed seemed to have grossed higher domestically than others. 1 each that passed and failed has grossed higher internationally than the others. There is 1 film that passed that has a very low metascore as compared to other films. 
```

## Scatterplot - Correlation Matrix

Write a paragraph discussing the output of the following

```{r, warning=FALSE, message=FALSE}
bechdel %>% 
  select(test, budget_2013, domgross_2013, intgross_2013, imdb_rating, metascore)%>% 
  ggpairs(aes(colour=test), alpha=0.2)+
  theme_bw()

# From the below output we can infer that there is a strong correlation between imdb rating and metascore among both movies that passed and failed the test. There also seems to be a correlation between the domestic gross and budget for the movie among both pass and fail movies.
```

## Categorical variables

Write a paragraph discussing the output of the following

```{r}
bechdel %>% 
  group_by(genre, test) %>%
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n))
 
bechdel %>% 
  group_by(rated, test) %>%
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n))

# Among the categories that seem to have a higher proportion of movies that fail are: Action with a 70/30 split, animation with an approx 70/30 split, crime with a 60/40 split, fantasy with a 70/30 split and mystery with a 75/25 split. In documentaries and sci-fi movies all the films have failed. In musical and thriller genres, all films pass. And in the horror genre there is a 70/30 split in favour of movies that have passed. In terms of ratings, NC-17 and G rated movies lean towards films that fail with an 85/15 and 60/40 split respectively.
```

# Train first models. `test ~ metascore + imdb_rating`

```{r}
lr_mod <- logistic_reg() %>% 
  set_engine(engine = "glm") %>% 
  set_mode("classification")

lr_mod
# Logistic regression model


tree_mod <- decision_tree() %>% 
  set_engine(engine = "C5.0") %>% 
  set_mode("classification")

tree_mod
#Decision tree model
```

```{r}

# Fitting the logistic regression model to the bechdel train data. test is the dependent variable and metascore and imdb are the independent variables.

lr_fit <- lr_mod %>% # parsnip model
  fit(test ~ metascore + imdb_rating, # a formula
    data = bechdel_train # dataframe
  )
# Fitting the decision tree model to the bechdel train data. test is the dependent variable and metascore and imdb are the independent variables.

tree_fit <- tree_mod %>% # parsnip model
  fit(test ~ metascore + imdb_rating, # a formula
    data = bechdel_train # dataframe
  )
```

## Logistic regression

```{r}
lr_fit %>%
  broom::tidy()
# This will output the information from the logistic regression model in a tidy format, it will return a data frame. 

lr_preds <- lr_fit %>%
  augment(new_data = bechdel_train) %>%
  mutate(.pred_match = if_else(test == .pred_class, 1, 0))
# This will generate predictions based on the bechdel train data. mutate adds a column that helps to evaluate correctness of the predictions. 

```

### Confusion matrix

```{r}
lr_preds %>% 
  conf_mat(truth = test, estimate = .pred_class) %>% 
  autoplot(type = "heatmap")

# Creates a heatmap of the confusion matrix. provides a visual output of the model in terms of its accuracy. 
```

## Decision Tree

```{r}
tree_preds <- tree_fit %>%
  augment(new_data = bechdel) %>%
  mutate(.pred_match = if_else(test == .pred_class, 1, 0)) 


```

```{r}
tree_preds %>% 
  conf_mat(truth = test, estimate = .pred_class) %>% 
  autoplot(type = "heatmap")
```

## Draw the decision tree

```{r}
draw_tree <- 
    rpart::rpart(
        test ~ metascore + imdb_rating,
        data = bechdel_train, # uses data that contains both birth weight and `low`
        control = rpart::rpart.control(maxdepth = 5, cp = 0, minsplit = 10)
    ) %>% 
    partykit::as.party()
plot(draw_tree)

```

# Cross Validation

Run the code below. What does it return?

```{r}
set.seed(123)
bechdel_folds <- vfold_cv(data = bechdel_train, 
                          v = 3, 
                          strata = test)
bechdel_folds

# Running code outputs a tibble that shows the data split into 10 cross-validation folds. 
```

## `fit_resamples()`

Trains and tests a resampled model.

```{r}
lr_fit <- lr_mod %>%
  fit_resamples(
    test ~ metascore + imdb_rating,
    resamples = bechdel_folds
  )


tree_fit <- tree_mod %>%
  fit_resamples(
    test ~ metascore + imdb_rating,
    resamples = bechdel_folds
  )
```

## `collect_metrics()`

Unnest the metrics column from a tidymodels `fit_resamples()`

```{r}

collect_metrics(lr_fit)
collect_metrics(tree_fit)


```

```{r}
tree_preds <- tree_mod %>% 
  fit_resamples(
    test ~ metascore + imdb_rating, 
    resamples = bechdel_folds,
    control = control_resamples(save_pred = TRUE) #<<
  )

# What does the data for ROC look like?
tree_preds %>% 
  collect_predictions() %>% 
  roc_curve(truth = test, .pred_Fail)  

# Draw the ROC
tree_preds %>% 
  collect_predictions() %>% 
  roc_curve(truth = test, .pred_Fail) %>% 
  autoplot()

```

# Build a better training set with `recipes`

## Preprocessing options

-   Encode categorical predictors
-   Center and scale variables
-   Handle class imbalance
-   Impute missing data
-   Perform dimensionality reduction
-   ... ...

## To build a recipe

1.  Start the `recipe()`
2.  Define the variables involved
3.  Describe **prep**rocessing [step-by-step]

## Collapse Some Categorical Levels

Do we have any `genre` with few observations? Assign genres that have less than 3% to a new category 'Other'

```{r}
#| echo = FALSE
bechdel %>% 
  count(genre) %>% 
  mutate(genre = fct_reorder(genre, n)) %>% 
  ggplot(aes(x = genre, 
             y = n)) +
  geom_col(alpha = .8) +
  coord_flip() +
  labs(x = NULL) +
  geom_hline(yintercept = (nrow(bechdel_train)*.03), lty = 3)+
  theme_light()
```

```{r}
movie_rec <-
  recipe(test ~ .,
         data = bechdel_train) %>%
  
  # Genres with less than 5% will be in a category 'Other'
    step_other(genre, threshold = .03) 
```

## Before recipe

```{r}
#| echo = FALSE
bechdel_train %>% 
  count(genre, sort = TRUE)
```

## After recipe

```{r}
movie_rec %>% 
  prep() %>% 
  bake(new_data = bechdel_train) %>% 
  count(genre, sort = TRUE)
```

## `step_dummy()`

Converts nominal data into numeric dummy variables

```{r}
#| results = "hide"
movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_dummy(all_nominal_predictors()) 

movie_rec 
```

## Let's think about the modelling

What if there were no films with `rated` NC-17 in the training data?

-   Will the model have a coefficient for `rated` NC-17? No
-   What will happen if the test data includes a film with `rated` NC-17? The model might end up ignoring the variable, or this might lead to less reliable predictions.

## `step_novel()`

Adds a catch-all level to a factor for any new values not encountered in model training, which lets R intelligently predict new levels in the test set.

```{r}

movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_novel(all_nominal_predictors) %>% # Use *before* `step_dummy()` so new level is dummified
  step_dummy(all_nominal_predictors()) 

```

## `step_zv()`

Intelligently handles zero variance variables (variables that contain only a single value)

```{r}
movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_novel(all_nominal(), -all_outcomes()) %>% # Use *before* `step_dummy()` so new level is dummified
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_zv(all_numeric(), -all_outcomes()) 
  
```

## `step_normalize()`

Centers then scales numeric variable (mean = 0, sd = 1)

```{r}
movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_novel(all_nominal(), -all_outcomes()) %>% # Use *before* `step_dummy()` so new level is dummified
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_zv(all_numeric(), -all_outcomes())  %>% 
  step_normalize(all_numeric()) 

```

## `step_corr()`

Removes highly correlated variables

```{r}
movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_novel(all_nominal(), -all_outcomes()) %>% # Use *before* `step_dummy()` so new level is dummified
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_zv(all_numeric(), -all_outcomes())  %>% 
  step_normalize(all_numeric()) 



movie_rec
```

# Define different models to fit

```{r}
## Model Building

# 1. Pick a `model type`
# 2. set the `engine`
# 3. Set the `mode`: regression or classification

# Logistic regression
log_spec <-  logistic_reg() %>%  # model type
  set_engine(engine = "glm") %>%  # model engine
  set_mode("classification") # model mode

# Show your model specification
log_spec

# Decision Tree
tree_spec <- decision_tree() %>%
  set_engine(engine = "C5.0") %>%
  set_mode("classification")

tree_spec

# Random Forest
library(ranger)

rf_spec <- 
  rand_forest() %>% 
  set_engine("ranger", importance = "impurity") %>% 
  set_mode("classification")


# Boosted tree (XGBoost)
library(xgboost)

xgb_spec <- 
  boost_tree() %>% 
  set_engine("xgboost") %>% 
  set_mode("classification") 

# K-nearest neighbour (k-NN)
knn_spec <- 
  nearest_neighbor(neighbors = 4) %>% # we can adjust the number of neighbors 
  set_engine("kknn") %>% 
  set_mode("classification") 
```

# Bundle recipe and model with `workflows`

```{r}
log_wflow <- # new workflow object
 workflow() %>% # use workflow function
 add_recipe(movie_rec) %>%   # use the new recipe
 add_model(log_spec)   # add your model spec

# show object
log_wflow


## A few more workflows

tree_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(tree_spec) 

rf_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(rf_spec) 

xgb_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(xgb_spec)

knn_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(knn_spec)

```

HEADS UP

1.  How many models have you specified?

    **Models defined: Logistic regression, decision tree, random forest, boosted tree, k-nearest neighbour.**

2.  What's the difference between a model specification and a workflow?

    **Model specification defines the structure of a machine learning model by specifying the variables and their relationships. A workflow uses the different functions to manipulate data to analyse and visualize data that we are working with.**

3.  Do you need to add a formula (e.g., `test ~ .`) if you have a recipe?

    **Yes, even when you have a recipe you need to add a formula to specify the variables and relationships between them example the target and predictor variables.**

# Model Comparison

You now have all your models. Adapt the code from slides `code-from-slides-CA-housing.R`, line 400 onwards to assess which model gives you the best classification.

```{r}
## Evaluate Models

## Logistic regression results{.smaller}

log_res <- log_wflow %>% 
  fit_resamples(
    resamples = bechdel_folds, 
    metrics = metric_set(
      recall, precision, f_meas, accuracy,
      kap, roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)) 

# Show average performance over all folds (note that we use log_res):
log_res %>%  collect_metrics(summarize = TRUE)

# Show performance for every single fold:
log_res %>%  collect_metrics(summarize = FALSE)



## `collect_predictions()` and get confusion matrix{.smaller}

log_pred <- log_res %>% collect_predictions()

log_pred %>%  conf_mat(test, .pred_class) 

log_pred %>% 
  conf_mat(test, .pred_class) %>% 
  autoplot(type = "mosaic") +
  geom_label(aes(
      x = (xmax + xmin) / 2, 
      y = (ymax + ymin) / 2, 
      label = c("TP", "FN", "FP", "TN")))


log_pred %>% 
  conf_mat(test, .pred_class) %>% 
  autoplot(type = "heatmap")


## ROC Curve

log_pred %>% 
  group_by(id) %>% # id contains our folds
  roc_curve(test, .pred_Pass) %>% 
  autoplot()


## Decision Tree results

tree_res <-
  tree_wflow %>% 
  fit_resamples(
    resamples = bechdel_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)
    ) 

tree_res %>%  collect_metrics(summarize = TRUE)


## Random Forest

rf_res <-
  rf_wflow %>% 
  fit_resamples(
    resamples = bechdel_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)
    ) 

rf_res %>%  collect_metrics(summarize = TRUE)

## Boosted tree - XGBoost

xgb_res <- 
  xgb_wflow %>% 
  fit_resamples(
    resamples = bechdel_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)
    ) 

xgb_res %>% collect_metrics(summarize = TRUE)

## K-nearest neighbour

knn_res <- 
  knn_wflow %>% 
  fit_resamples(
    resamples = bechdel_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)
    ) 

knn_res %>% collect_metrics(summarize = TRUE)


## Model Comparison

log_metrics <- 
  log_res %>% 
  collect_metrics(summarise = TRUE) %>%
  # add the name of the model to every row
  mutate(model = "Logistic Regression") 

tree_metrics <- 
  tree_res %>% 
  collect_metrics(summarise = TRUE) %>%
  mutate(model = "Decision Tree")

rf_metrics <- 
  rf_res %>% 
  collect_metrics(summarise = TRUE) %>%
  mutate(model = "Random Forest")

xgb_metrics <- 
  xgb_res %>% 
  collect_metrics(summarise = TRUE) %>%
  mutate(model = "XGBoost")

knn_metrics <- 
  knn_res %>% 
  collect_metrics(summarise = TRUE) %>%
  mutate(model = "Knn")

# create dataframe with all models
model_compare <- bind_rows(log_metrics,
                           tree_metrics,
                           rf_metrics,
                           xgb_metrics,
                           knn_metrics) 

#Pivot wider to create barplot
  model_comp <- model_compare %>% 
  select(model, .metric, mean, std_err) %>% 
  pivot_wider(names_from = .metric, values_from = c(mean, std_err)) 

# show mean are under the curve (ROC-AUC) for every model
model_comp %>% 
  arrange(mean_roc_auc) %>% 
  mutate(model = fct_reorder(model, mean_roc_auc)) %>% # order results
  ggplot(aes(model, mean_roc_auc, fill=model)) +
  geom_col() +
  coord_flip() +
  scale_fill_brewer(palette = "Blues") +
   geom_text(
     size = 3,
     aes(label = round(mean_roc_auc, 2), 
         y = mean_roc_auc + 0.08),
     vjust = 1
  )+
  theme_light()+
  theme(legend.position = "none")+
  labs(y = NULL)

## `last_fit()` on test set

# - `last_fit()`  fits a model to the whole training data and evaluates it on the test set. 
# - provide the workflow object of the best model as well as the data split object (not the training data). 
 
last_fit_xgb <- last_fit(xgb_wflow, 
                        split = data_split,
                        metrics = metric_set(
                          accuracy, f_meas, kap, precision,
                          recall, roc_auc, sens, spec))

last_fit_xgb %>% collect_metrics(summarize = TRUE)

#Compare to training
xgb_res %>% collect_metrics(summarize = TRUE)


## Variable importance using `{vip}` package

library(vip)

last_fit_xgb %>% 
  pluck(".workflow", 1) %>%   
  pull_workflow_fit() %>% 
  vip(num_features = 10) +
  theme_light()


## Final Confusion Matrix

last_fit_xgb %>%
  collect_predictions() %>% 
  conf_mat(test, .pred_class) %>% 
  autoplot(type = "heatmap")


## Final ROC curve
last_fit_xgb %>% 
  collect_predictions() %>% 
  roc_curve(test, .pred_Pass) %>% 
  autoplot()

# Looking at the output comparing each of the models, the Random Forest model seems the give the best classification. Out of all the models, it has a 67% accuracy that is the highest. Close to the Random Forest is the XGBoost model with a 63% accuracy. The lowest performing model seems to be the Logistic Regression with an accuracy of 47%. 
```

# Details

-   Who did you collaborate with: Jenna Thomas
-   Approximately how much time did you spend on this problem set: 3-4 hours
-   What, if anything, gave you the most trouble: To figure out how to adapt the code from the CA housing. It also took me some time to understand all the other code written throughout the file to see which code was outputting which model.
